### 2015-03-31 casestudies.tex
some work

\section{Case Study\label{sec:casestudy}: visual exploration of topic modeling in RCloud}

As an example, we present a simple real-world application
deployed under RCloud.

The application is an implementation of LDAVis,
a recent method for visualizing large text corpora.
It was designed to help non-experts to explore
collections of short text documents using topic modeling and
visualization. Topic modeling~\cite{Blei:2003:LDA} is a standard
technique for text summarization which, although powerful, requires
some manual intervention and interpretation~\cite{Sievert:2014:LAM}.

\begin{figure*}
  \includegraphics[width=\linewidth]{fig/casestudytext/casestudytext.pdf}
  \caption{\label{fig:textvis}An example application developed and deployed in RCloud.}
\end{figure*}

LDAVis was developed by two technical staff members at
AT\&T Labs, and originally was written in RStudio's
Shiny~\cite{RStudio:2013:SWA}, a framework for developing R
applications for the web.  While Shiny provides outstanding ease of
development, it turns out that discoverability and deployment are equally
important aspects in the lifecycle of an internal application (see
Section~\ref{sec:interviews}). In these aspects, RCloud provides a
simple solution: \emph{all developed notebooks are automatically deployed}.

LDAVis combines text analysis, dimensionality reduction and
interactive visualization. Text analysis is performed by
combining a standard R library to fit LDA models using
Gibbs sampling~\cite{} with custom R code written by the
developers. The analysis module is, ultimately, a single R function
exposed to the web application via the Javascript-R RPC
mechanism described in Section~\ref{sec:system};
thus analysis is performed remotely on an RCloud server.

Each text topic is a probability distribution over the all the
words of the document. To expose patterns in the relationships
between topics, LDAVis employs a combination of
interactive visualization and dimensionality reduction,
allowing users to adjust measures for topic distances and the choice
of dimensionality reduction technique. The dimensionality reduction
algorithms and distance measures are implemented in R, which
means they can be executed on the RCloud servers as well.

The result of the dimensionality reduction process is a
two-dimensional plot of the topic space, an example of which is shown in
Figure~\ref{fig:textvis}. The interactive view is implemented in SVG
and Javascript through D3~\cite{Bostock:2011:DDD}. One of the most
popular web-based visualization libraries for R is ggvis~\cite{ggvis},
so it is natural to ask if ggvis could have been used instead of
custom Javascript. In this case, the interactive features of ggvis
(and, by extension, Shiny) are a subset of those in Vega~\cite{vega}.
Custom interactions in LDAVis (such as hovering over a topic,
topic cluster, or word) do not appear to be available yet in Vega~\cite{vega},
although the required components for reactive interaction were
recently described by Satyanarayan et
al.~\cite{Satyanarayan:2014:DID}.
Although custom Javascript was required, the flexibility is welcomed
by many web developers.

The LDAVis application is simple, but highlights some unique features
of RCloud. While RCloud notebooks allow deployment of R analyses over
the web with no additional effort, RCloud \emph{applications} are more
powerful, and are developed with a combination of Javascript and HTML
for the front end. This requires additional knowledge over Shiny, but
we argue that the RCloud model makes the analysis side simpler 
(since analysts simply write R in the style they are used to),
and the front end visualization side is simpler for the web
developers (since they simply write the Javascript code in
the style \emph{they} are used to). In addition, RCloud applications
inherit the automatic deployment and discoverability features of
all RCloud notebooks.

%% Text analysis in

%% To illustrate some aspects of RCloud, we present an example notebook
%% for stock price analysis.\footnote{Unfortunately this report could not
%% include production notebooks containing proprietary information.}
%% The example, although simplified, shows key steps in the development
%% and deployment of an application.

%% \begin{figure}
%% \includegraphics[width=\linewidth]{fig/casestudy1/casestudy1.pdf}
%% \caption{\label{fig:stockvis}Iterations of a stock ticker
%%   visualization, based on an example by Hadley Wickham. A
%%   simple, static visualization of the closing price of a single stock
%%   is progressively developed into a configurable display suitable
%%   for a dashboard, then into an interactive visualization that
%%   compares the volatility and volume of two stocks, and finally
%%   into an API for data access by other RCloud notebooks. The notebooks
%%   in this example can all be loaded as web pages. When a notebook
%%   corresponding to a function call is displayed a web page,
%%   its associated documentation is displayed.}
%% \end{figure}

%% It shows a sequence of visualizations of the performance
%% of financial stocks over several years. The first visualization
%% in this example uses ggplot2 and was written by Hadley Wickham.
%% It reads data provided by Yahoo! Finance as a web service, and shows
%% the price for a single trading symbol.

%% The webpage that produces that visualization includes a link to
%% the underlying source code. From this link, a user can fork the notebook.
%% In the example,a configurable ticker is added, based on the URL of
%% the notebook. The change to the notebook is minor.

%% Next, the notebook author or a collaborator may decide to extend
%% the application to also provide convenient access to the data,
%% apart from its visualization.
%% This is done by creating a notebook that defines a function.
%% That notebook then becomes a \emph{subroutine} for other notebooks.
%% It can be invoked from interactive notebooks, such as dashboards.
%% The data access notebook is version-controlled like other notebooks.

%% Finally, consider the situation where an analyst wants to understand
%% the price dynamics of stocks with respects to other attributes and
%% time ranges. For this, interactive visualization may be very helpful.
%% In our example, the analyst creates an interactive tool with multiple
%% linked views.

%% Demonstrating the integration of R and JavaScript, here the analyst
%% writes an R function that passes the dataframe and description of
%% the charts to the dcplot library.  Simple R expressions
%% are captured as trees to generate JavaScript expressions.
%% The terse chart description language, with sensible defaults inspired by
%% ggplot2, provides a simple yet powerful interface to the grouping
%% and reduction functionality of the well-accepted charting libraries
%% crossfilter, dc.js and d3.js.

% \subsection{Text analysis\label{sec:textvis}}
% Same.

% IMPORTANT: what is unique about RCloud here
% From prototyping to dashboard
% Getting other information from the web
%