### 2015-06-16 conclusion.tex
yes

\section{Conclusions and Future Work}
We designed a prototype environment for exploration, sharing,
presentation, and publishing.
This study has provided a step toward practical ``DevOps for data science''
and reproducible, publishable experiments. Our users found that the
benefits of sharing

We found that features for sharing and publishing were eagerly adopted.
Features for single-user data exploration, that compete with existing mature
tools, were not accepted as readily. Some experienced users
fashioned their own workflow so they could keep using familiar EDA tools.

Possible next steps to improve the existing prototype are to incorporate
richer recommendation techniques, to provide fine-grained information
security, and to improve the usability of the human interface.

There is a clear need for better software infrastructure to support
experiments in visual analytics systems. We did benefit greatly from
building RCloud on exisitng software and services, but the implementation
was still difficult with much trial and error. Possibly some analogue
of grammar-based visualization methods would help the visual analytics process.

RCloud code is available at \url{github.com/att/rcloud/}
under an MIT open source license.