### 2015-06-23 conclusion.tex
reduced text in conclusion.

\section{Conclusions and Future Work}
We designed a prototype environment for exploration, sharing,
presentation, and publishing.  This study has provided a step toward
practical ``DevOps for data science'' and reproducible, publishable
experiments.

RCloud has become a valid avenue for easy deployment of visualizations
and interactive exploration tools inside our organization.  We found
that features for sharing and publishing were eagerly adopted.
Features for single-user data exploration, on the other hand, competed
with existing mature tools, and were not accepted as readily.  In
future work, we hope to study and to understand the degree to which
this reluctance is caused by bad design decisions on our part, and how
much comes from ``mere'' change aversion.

We still firmly believe that the current biggest barrier to the
adoption of visual analytics is in software infrastructure.
Just like structuring visualization software around
simple but composable parts is a successful approach (like Vega, ggplot and the grammar of
graphics), it is intriguing to consider what systems are
possible if this philosophy were carried over to the 
broader requirements of interactive analysis.

RCloud is open-source, and available at \url{github.com/att/rcloud/}.