### 2015-06-27 revision_letter.tex

\documentclass{article}
\usepackage{microtype}
\usepackage{color}
\title{Letter to Reviewers of Submission \#341}
\begin{document}
\maketitle

We thank the reviewers for their careful reading of the manuscript and
their suggestions. In this document, we summarize the changes we
performed to the original submission.

\paragraph*{Paper Length} We have reduced the length of the paper
to eight pages, removing only extended discussion in the limitation
and interviews sections. We achieved most of the space savings was
achieved by careful redesign of the figures and tables, and rewriting
of text.

\paragraph*{Section Lengths} The reviewers were concerned about the length
of the introduction. We have significantly reduced it.

\end{document}